# Introduction

In this chapter we will see the different forms of codelists (`<codelist>`, `<codelist_with_details>` and `<concept_set_expression>`) and we will create some custom codelists and see how to change between them.

For this chapter we will use mainly the [CodelistGenerator]() package and [AthenaR]() to create a reference to the vocabulary database. In this case we will use the OMOP CDM vocabularies of **February 2026**(v20260227).

```{r, eval=FALSE}
library(CodelistGenerator)
library(AthenaR)
library(dplyr)

cdm <- cdmFromVocabulary(vocabulary = "v20260227", source = "duckdb")
```

```{r, echo=FALSE}
library(CodelistGenerator)
library(dplyr)

cdm <- omock::mockCdmFromDataset(source = "duckdb")
```


## `<codelist>`

A `<codelist>` is a named list of codes, we will create codelists with the function `newCodelist()` from the [omopgenerics]() package (omopgenerics is the package that defines all the classes of the objects we use).

```{r}
codelist <- newCodelist(list(tubercusosis = c(1L, 2L, 3L), covid_19 = c(4L, 5L)))
codelist
```

we can easily validate that the codes exists in the cdm using the cdm argument:

```{r}
codelist <- newCodelist(x = list(tubercusosis = c(1L, 2L, 3L), covid_19 = c(4L, 5L)), cdm = cdm)
codelist
```

## `<codelist_with_details>`
